Embodied Neuromorphic Benchmarks

### 2015-08-15 paper/paper.tex
Added numbers for power comparison

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing [TODO: ref??].  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
functional well when deployed in real situations.

@@

These considerations suggest that there is a neuromorphic-friendly family of 
algorithms to address the general problem of controlling a wide variety of 
physical systems.  Identifying those algorithms will allow us to benchmark 
their performance across example tasks and physical configurations.  
To implement these algorithms, the input to the
neuromorphic hardware is $q$, the system state.  This input is fed to each
neuron such that each neuron produces some output activity that is based on
this input.  Since $q$ will be multi-dimensional (if there is more than one
joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q + b_i$, 
where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen 
vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities,
or could be learned using some back-propogation of error method.  Here, for simplicity, we only
consider the approach of randomly selecting $e_i$ and $b_i$.},
and $b_i$ is a randomly chosen bias term).  
Given this input, the neurons will produce some output $A$.
We now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
that is initially all zeros.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
However, on that benchmark all three systems are implementing exactly 500
neurons.  The only thing that can be concluded from this is that the
differences in neuron model across that hardware does not significantly
impact performance.  A more interesting measure would be to compare based
on how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  In this case, the CPU can
run XXX neurons, the GPU XXX neurons, and a single SpiNNaker chip can run XXX
neurons. Furthermore, we also show the effects of increasing the number
of joints in the simulation.

@@

While it is possible to run large neural models on standard CPUs and GPUs,
one of the primary advantages of neuromorphic hardware is its power
efficiency.  For this reason, the third benchmark fixes the number of neurons
at however many that hardware can compute in 0.5 Watts.  While this is a
reasonable number for SpiNNaker, neither the CPU nor the GPU can even run
with that little power.  For this reason, we simply take the number of
neurons that hardware can simulate in real time and scale it by the power
consumption of the hardware.  The result is that the CPU can run XXX neurons,
the GPU XXX neurons, and SpiNNaker can run XXX neurons.  The results of
this benchmark are shown in Figure \ref{fig:analysis_power}.  This clearly
shows the power advantage of neuromorphic hardware over standard CPUs and
GPUs on this task.

@@

The benchmarks described above all use the same underlying minimal simulation
as a way to characterize the overall performance of particular hardware
across a range of situations.  By adjusting the random distributions that
define that range of situations, we generate different benchmarks that
explore the capabilities of the systems in different ways.  This allows for
an explicit depiction of the sorts of conditions in which particular
neuromorphic hardware perfroms well.  After all, it is unlikely that one
piece of neuromorphic hardware will be the best choice in all situations;
rather, these benchmarks allow us to demonstrate the advantages and disadvantages
of the hardware by looking at the same underlying system, but with multiple
different distributions of parameters.

Python software for the minimal simulation and the full benchmarks are
available at \texttt{http://github.com/ctn-waterloo/ctn_benchmarks}.

@@

However, it is also clear that performance these benchmarks is a result of
a combination of the hardware itself, the algorithm being run, and the system that
interfaces the hardware to the environment.  Thus, for any given hardware, 
we can explore improvements to the algorithm (better choices for $e$, 
different learning rules, adaptive learning rates, adapting $K_p$ and $K_d$, etc.).
For example, in the SpiNNaker hardware implementation not only can the neuron model
be adjusted, but the distribution of the task across the multiple cores is also
under programmer control.  Futhermore, SpiNNaker provides a custom I/O interface
for high-speed communication that could be used to reduce communication delay.

@@

The most straightforward expansion would be to consider motor control where
the state being controlled \emph{is not the motor state itself}.  That is,
suppose we want to control the position of a hand $x$, but our output $u$
only directly controls the joints $q$ of an arm.  The position of the hand $x$
is some function of $q$, but this function may be unknown or highly complex.
Crucially, there is a learning rule that is similar to the one discussed above
that can learn this mapping (\citealt{adaptiveJacobian}), and we have had
some success in using it for particular arm control tasks (citealt{dewolf2014}).
This would give a benchmark with a much more flexible set of capabilities;
for example a non-neural version of this learning algorithm was able to
learn to control a robot arm to follow a certain path, given only the
visual feedback of where the tip of its hand was, without knowing the
geometry of its own hand or the position of the camera with relation to that
hand (\citealt{adaptiveJacobian}).  While it is unlikely that such a
system would work for \emph{any} possible sensorimotor combination, developing
such a benchmark could identify the space of such embodiments that a
particular algorithm running on particular neuromorphic hardware would be
good at controlling.

>>>>

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing 
(see \citealt{Jakobi95}; \citealt{Koos2013}).  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
functional well when deployed in real situations.

@@

These considerations suggest that there is a neuromorphic-friendly family of 
algorithms to address the general problem of controlling a wide variety of 
physical systems.  Identifying those algorithms will allow us to benchmark 
their performance across example tasks and physical configurations.  
To implement these algorithms, the input to the
neuromorphic hardware is $q$, the system state.  This input is fed to each
neuron such that each neuron produces some output activity that is based on
this input.  Since $q$ will be multi-dimensional (if there is more than one
joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q + b_i$, 
where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen 
vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities,
or could be learned using some back-propagation of error method.  Here, for simplicity, we only
consider the approach of randomly selecting $e_i$ and $b_i$.},
and $b_i$ is a randomly chosen bias term).  
Given this input, the neurons will produce some output $A$.
We now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
that is initially all zeros.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
However, on that benchmark all three systems are implementing exactly 500
neurons.  The only thing that can be concluded from this is that the
differences in neuron model across that hardware does not significantly
impact performance.  A more interesting measure would be to compare based
on how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  In this case, with the
current implementation, the CPU can
run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
there are 18 on one chip) can run 500
neurons.  These values were measured empirically with the current versions
of the reference Nengo implementation, the Nengo OpenCL implementation,
and the Nengo SpiNNaker implementation (as of August 10, 2015).  
Figure \ref{fig:analysis_compute}
also shows the effects of varying the number of joints in the simulation,
giving an indication of the complexity of the external forces that can be
learned by this many neurons.

@@

While it is possible to run large neural models on standard CPUs and GPUs,
one of the primary advantages of neuromorphic hardware is its power
efficiency.  For this reason, the third benchmark normalizes the number of neurons
based on power consumption.  With a power budget of 1W per chip (with 16 used cores),
we estimate 0.0625W for the 500 neurons used here and round up to 0.1W to be
conservative.  Of course, neither the CPU nor the
GPU are designed to run on that little power.  For this reason,
on this benchmark we simply and scale it by the power consumption of the hardware.
For this power consumption we measure the difference between the idle power
consumption and the consumption when running the benchmark.  For the Intel
i5-3337U this was $34W - 11.5W = 12.5W$ and for the Nvidia Tesla C2075 GPU
this was $74W - 70W = 4W$.  This value is much lower than the peak power
consumption supported by the GPUi ($215W$), indicating that the current implementation
does not make extensive use of the GPU.  Indeed, initial analysis indicates that
the main bottleneck is communication between the GPU and the environment (i.e.,
the minimal simulation), and we feel it is appropriate that this benchmark
captures that limitation of the current GPU implementation.

Given the above considerations, we run the benchmark indicating that the CPU can run 42 neurons per 0.1W,
the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  If
this number of neurons is kept fixed and the number of joints controlled is
varied, we get the benchmark shown in Figure \ref{fig:analysis_power}.  This clearly
shows the power advantage of neuromorphic hardware over standard CPUs and
GPUs on this task.

@@

The benchmarks described above all use the same underlying minimal simulation
as a way to characterize the overall performance of particular hardware
across a range of situations.  By adjusting the random distributions that
define that range of situations, we generate different benchmarks that
explore the capabilities of the systems in different ways.  This allows for
an explicit depiction of the sorts of conditions in which particular
neuromorphic hardware performs well.  After all, it is unlikely that one
piece of neuromorphic hardware will be the best choice in all situations;
rather, these benchmarks allow us to demonstrate the advantages and disadvantages
of the hardware by looking at the same underlying system, but with multiple
different distributions of parameters.

Python software for the minimal simulation and the full benchmarks are
available at \texttt{http://github.com/ctn-waterloo/ctn\_benchmarks}.

@@

However, it is also clear that performance these benchmarks is a result of
a combination of the hardware itself, the algorithm being run, and the system that
interfaces the hardware to the environment.  Thus, for any given hardware, 
we can explore improvements to the algorithm (better choices for $e$, 
different learning rules, adaptive learning rates, adapting $K_p$ and $K_d$, etc.).
For example, in the SpiNNaker hardware implementation not only can the neuron model
be adjusted, but the distribution of the task across the multiple cores is also
under programmer control.  Furthermore, SpiNNaker provides a custom I/O interface
for high-speed communication that could be used to reduce communication delay.

@@

The most straightforward expansion would be to consider motor control where
the state being controlled \emph{is not the motor state itself}.  That is,
suppose we want to control the position of a hand $x$, but our output $u$
only directly controls the joints $q$ of an arm.  The position of the hand $x$
is some function of $q$, but this function may be unknown or highly complex.
Crucially, there is a learning rule that is similar to the one discussed above
that can learn this mapping (\citealt{adaptiveJacobian}), and we have had
some success in using it for particular arm control tasks (\citealt{dewolf2014}).
This would give a benchmark with a much more flexible set of capabilities;
for example a non-neural version of this learning algorithm was able to
learn to control a robot arm to follow a certain path, given only the
visual feedback of where the tip of its hand was, without knowing the
geometry of its own hand or the position of the camera with relation to that
hand (\citealt{adaptiveJacobian}).  While it is unlikely that such a
system would work for \emph{any} possible sensorimotor combination, developing
such a benchmark could identify the space of such embodiments that a
particular algorithm running on particular neuromorphic hardware would be
good at controlling.
